---
title: "ST404 Assignment 2"
author: "Frank Or, Remos Gong, Sam Glanfield, Thomas Broadbent"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2: default
  pdf_document: 
    number_sections: true
fontsize: 11pt
linestretch: 1.5
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.align = "center",
	fig.width = 7,
  echo = TRUE)
library(bookdown)
library(knitr)
```

```{r libraries, include=FALSE}
library(dplyr)
library(car)
library(tidyr)
library(glmnet)
```

\pagebreak 

# Findings

## Summary of EDA
In the previous EDA report we have explored the data set and we will perform data cleaning and variable transformations before constructing a linear model. 

---Might Merge in one if run out of space?

### Missing/Incorrect Values and Outliers

We identified 152 missing values in __Percent Employed 16 and Over__. Since they have no pattern, we deduced that these values are missing at random (MCAR). We used other complete data entries to calculate what we would expect these values to be and impute them back to our data set. We also identified values in __Average Household Size__ that unreasonably small. We believed this is an error in data entry and we scaled them by 100 in order to make it normal.
For outliers, We removed __Williamsburg City, Virginia__ because of its high __Incidence Rate__ but low __Death Rate__, hence high influence to our model. 

### Transformations

In order to fulfill the model assumptions, namely __Linearity__, __Homoscedasticity__, __Normality__, we performed log transforms for two variables: __Median Income__ and __Percent Black__ which suffered the most among all variables. Although the transformations did not cure the problems, they result in improvements in linearity and normality, reducing heteroscedasticity for both said variables. For other variables, we deduced that they are good enough to fulfill the model assumptions and hence did not perform any transformations for model simplicity.

--Might add plots?

### Multicollinearity

We have discovered a few pairs or clusters of predictor variables that are highly correlated with each other. The obvious ones are __Median Age Female__ and __Median Age Male__, __Percent Married__ and __Percent Married Households__, __Binned Income__ and __Median Income__. These variables measure the same force and hence one of the pairs can be dropped in early modelling. The others such as the employment variables and the heath coverage variables have less evidence of collinearity/ multicollinearity and are hence included in the modelling.

### Correlation with Death Rate

While most predictor variables have absolute correlation coefficient roughly 0.3 to 0.4 with death rate, Median Age Male, Median Age Female and Average Household Size have almost zero correlation coefficient. We expect our model to not include these uncorrelated variables.

## Major Determinants of Mortality Rates

From our model building process we have identified that the major determinants in high mortality rates of cancer in the US are: __Incidence Rate, Percent Unemployed 16 and Over, Percent Employer Provided Private Coverage and Percent Black__. Whereas, the major determinants in low mortality rates are: __Percent Employed 16 and Over, Percent Private Coverage, Education Levels and Median Income__.

## Modelling Approach

We used Stepwise Regression, RIDGE Regression and LASSO Regression to build our model. We compared the outputs and the variables selections of these models. We also analyse the goodness of fits of these models using Leave-one-out cross validation, R-squared statistics and residuals analysis.
We did not include _Geography_ and __Binned Income__ and we were using the _Log Percent Black__ and __Log Median Income__ in the following analysis.

---I found the following sections very tricky to write up. Might need a better structure. NEED HELP!

### Stepwise Regression
We used both __Akaike’s Information Criteria(AIC)__ and __Bayesian Information Criteria(BIC)__ in our analysis. The latter one penalises additional parameters harder. We performed forward, backward and hybrid stepwise regression. In all of the models that were generated we notice that we have groups of parameters that are similar and hence induce multicollinearity in the models. For example, we see both __Percent Married__ and __Percent Married Households__ in the models and as stepwise methods do not account for multicollinearity this causes increases in the variance of our coefficient estimates and makes the model sensitive to changes, thus reducing drastically the predictive power of the model. In order to address the issue of multicollinearity we apply Ridge and LASSO techniques which account for this.

### RIDGE Regression

### LASSO Regression

The suggested model contains eight variables which is the simplest model among our approaches. Since the purpose of this report is to reveal patterns in the mortality rate, we believed it is best to choose a simple model for stronger explanatory power.

### Final Model Choice and Diagnostics

We used residuals plot and QQ plot to diagnose the stepwise regression model. Both plots are satisfactory, agreeing with model assumptions.

```{r statestepwise Findings, echo=FALSE, fig.asp=0.5}
# Residual plots of state stepwise regression
par(mfrow=c(1,2))
plot(hybridoptimalBIC2, which=c(1,2))
```

We also performed Leave-one-out cross-validation for this stepwise model, which gives $R^2$ value of $0.4621$.
We did the same cross-validation for our Ridge and Lasso model.
The Ridge model gives $R^2$ value of $0.4436$.
The Lasso model gives $R^2$ value of $0.4441$.

The stepwise model fits the observed data the best among all our models. The Lasso and the Ridge model is similar in this sense.
However, the Lasso model contains the least number of predictor variables and we believed this is the most important criterion.
Thus from our analysis briefly outlined above and in the next section we recommend the following model, as seen below, due to its mix of predictive and explanatory power as well as being a simple model.

$DeathRate = 470.47 + 0.23IncidenceRate - 0.25PctEmployed16\_Over + 0.029PctUnemployed16\_Over - 0.32PctPrivateCoverage + 0.21PctEmpPrivCoverage - 10.06Edu18\_24 + 0.90log(PctBlack) - 32.34log(MedianIncome)$

## Areas that do not conform the general pattern

We do see examples of counties that do not conform to the general pattern with unusually high or low mortality rates. One example we have is 'Williamsburg City, Virginia' where we see a high incidence rate of 1014 yet a comparatively small death rate of 162, however we see that it's percent private coverage is 19% above the US average and the median male and female ages of 26 and 24 respectively are considerably lower than the median US ages, potentially being causes for this low death rate. Similarly, we see from the average residual map that in particular the 3 states Utah (Average of -22), Idaho (Average of -19) and Colorado (Average of -18) all have high residual averages and as they are negative this implies that the model is overfitting for these states and thus these states don't conform to the general pattern. However, as the model if overfitting and we are estimating death rate this isnt a cause for concern as we know the average death rates for these states is highly likely to be lower than what the model predicts. It is also important to note that these all lie in a similar region and in general we see that the majority of states on the west coast have an average negative residual, so the model overfits for these states, hwereas on the east coast and central we see positive average residuals, the model is underfitting these states. With the main staes of concern for overfitting with an average residual of 14 each for Oklahoma and Arkansas and an average of 13 in the District of Columbia.

# Statistical Methodology

```{r loadData, include=FALSE}
## Included Libraries
# For pipe operator and general mutation
load('cancer.rdata')
```

We first combined our results and findings from our preliminary EDA which is mainly discussed in [Outliers](#outliers) and [Transformations](#transformations).

## Outcomes of EDA

### Missing Or Incorrect Values

We also see counties with missing values in Percentage Employed 16 and Over and we conclude that the data is Missing Completely at Random. In order to rectify this we impute this data by fitting a linear regression model of Percentage Employed 16 and Over on the remaining variables to estimate what these values would be.

```{r impute, include=FALSE}
# Impute the missing data seen in the data set
mod1=lm(PctEmployed16_Over~+incidenceRate+medIncome+binnedInc+povertyPercent+MedianAgeMale+MedianAgeFemale+AvgHouseholdSize+PercentMarried+PctUnemployed16_Over+PctPrivateCoverage+PctEmpPrivCoverage+PctPublicCoverage+PctBlack+PctMarriedHouseholds+Edu18_24,cancer)
missdf = cancer[which(is.na(cancer$PctEmployed16_Over)==TRUE),]
imputed = predict(mod1,missdf)
cancer[which(is.na(cancer$PctEmployed16_Over)==TRUE),"PctEmployed16_Over"] = imputed
```

For counties with Average Household Size less than one we took the decision to scale the transformations by 100 and keep them in the dataset. This fixed the normality of AvgHouseholdSize as shown in the histogram.

```{r averageHouseholdSize, include=FALSE}
# Scale average household sizes that are less than 1 by 100
cancer[which(cancer$AvgHouseholdSize < 1), ]$AvgHouseholdSize <- 
  100*cancer[which(cancer$AvgHouseholdSize < 1), ]$AvgHouseholdSize
hist(cancer$AvgHouseholdSize, breaks=30, xlab="AvgHouseholdSize", main="Histogram of AvgHouseholdSize")
```

### Outliers {#outliers}

Counties with high Incidence Rates, namely 'Union County, Florida' and 'Williamsburg City, Virginia'. We looked into the cook's distance plots and noticed only 'Williamsburg City, Virginia' has large cook's distance and hence influential. The first cook's distance plot used a linear model with only incidenceRate as the predictor variable. The second used all the numerical variables. We concluded although 'Union County, Florida' has high leverage, it is not influential and hence should be kept in our data set.

```{r incidenceRateRemoval, echo=FALSE, fig.asp=0.5, fig.cap="Cook's distance plots with one variable(Left) and all variables(Right)"}
# Cook's distance Plot
par(mfrow=c(1,2))
plot(lm(deathRate ~ incidenceRate,data=cancer),4)
plot(lm(deathRate ~ .,data=cancer[-c(1,4)]),4)
# Removing outlier incidence rates 'Williamsburg City, Virginia'
cancer <- filter(cancer, incidenceRate <= 850)
```

### Transformations {#transformations}

We transform Percent Black by first shifting the values upwards by 0.05, to ensure we have no zero values, then take a log transform. We also transform the Median Income by again taking a log transformation. We do these transformations to ensure the data is not heavily skewed and allow for a more accurate model.

```{r transformations, include=FALSE}
# Log transforming the heavily skewed distributions of PctBlack and medIncome
cancer$logpctblack = log(cancer$PctBlack+0.05)
cancer$logmedincome = log(cancer$medIncome)
```

The following residual plots show the improvements in homoscedasticity in PctBlack and medIncome after log-transform respectively.

```{r heteroscedasticity1,echo=FALSE, fig.asp=0.5, fig.cap="Residuals plot Percent Black and Log Percent Black"}
# Showing improvements in homoscedasticity in PctBlack
par(mfrow=c(1,2))
plot(lm(deathRate~PctBlack,data=cancer),1)
plot(lm(deathRate~logpctblack,data=cancer),1)
```
---Might add other improvements 

## Modelling Approach and Variable Selection

### AIC and BIC Forward and Backward Variable Selection

We perform forward, backward and hybrid stepwise regression according to both AIC and BIC. We see that the models generated for AIC are all the same and the ones generated for BIC are all the same. In the AIC case we have the model has 12 parameters, whereas the BIC model has 11 variables. 

```{r stepwise, include=FALSE}
# Below we perform stepwise regression for both AIC and BIC
# cancermodel = cancer[,-c(1,3,15)]
cancermodel <- cancer %>% select(
  !c("Geography", "medIncome", "binnedInc", "PctBlack"))
c0=lm(deathRate~1,cancermodel)
cmax=lm(deathRate~.,cancermodel)
forwardoptimalAIC = step(c0,direction="forward",
scope=list("lower"=c0,"upper"=cmax),trace=0)
backwardoptimalAIC = step(cmax,direction="backward",
scope=list("lower"=c0,"upper"=cmax),trace=0)
hybridoptimalAIC = step(c0,direction="both",
scope=list("lower"=c0,"upper"=cmax),trace=0)
forwardoptimalBIC = step(c0,direction="forward",
scope=list("lower"=c0,"upper"=cmax),trace=0,k=log(3044))
backwardoptimalBIC = step(cmax,direction="backward",
scope=list("lower"=c0,"upper"=cmax),trace=0,k=log(3044))
hybridoptimalBIC = step(c0,direction="both",
scope=list("lower"=c0,"upper"=cmax),trace=0,k=log(3044))
```

From the summary output we see that for the AIC model all coefficients have strong evidence that they are different from zero other than Percent Unemployed 16 and Over and Percent Public Coverage. For the BIC model these variables are not in the model and further from the summary output we see with strong evidence that all the coefficients are different from zero. From observing the output we see that we would expect there to be multicollinearity in these models due to variables that measure similar or opposite quantities, for example Percent Married and Percent Married Households.

```{r summaryAICBIC, include=FALSE}
# We compute the summaries of the stepwise models
summary(hybridoptimalAIC)
summary(hybridoptimalBIC)
```

Below we see the residual plots for the stepwise BIC model and can see in the left plot that we appear to have constant variance and the residuals have a mean of zero satisfying these assumptions. 

```{r stepwise plots,echo=FALSE,fig.asp=0.5, fig.cap="Residuals plot of Hybrid Stepwise Model using BIC"}
plot(hybridoptimalBIC,which=1)
```


We compute the Variance Inflation Factors (VIF) for the BIC model and see that there are numerous values that are at least 5 indicating that we have multicollinearity. Specifically, we see a VIF of at least 10 in Percent Married and a VIF of 8 in Percent Married Households. As well as in log(Median Income), Percent Employed 16 and Over, Percent Employers Private Coverage and Percent Private Coverage. Due to this frequent multicollinearity between the predictors in the stepwise model this suggests that the stepwise model is not a suitable approach as multicollinearity is not taken into account and thus we proceed below with stepwise regression with the states where we encounter a similar problem and further to Ridge Regression and Lasso to reduce the effects of multicollinearity on the model.

```{r StepwiseVIF, include=FALSE}
# Computing the VIF of the BIC stepwise regression model
vif(hybridoptimalBIC)
```

Now we perform stepwise with states included as a factor and as can be seen in both the AIC and BIC case the factor State is kept in the model and otherwise the other variables included are similar with the main different exclusion being log(Percent Black) which was included in the model without the addition of the states.

```{r statestepwise, include=FALSE}
# Create stepwise regression models including the states
cancermodel2 = separate(cancer,"Geography", into=c("County","State"),sep=",")[-c(1,4,5,16)]
c0=lm(deathRate~1,cancermodel2)
cmax=lm(deathRate~.,cancermodel2)
hybridoptimalAIC2 = step(c0,direction="both",
scope=list("lower"=c0,"upper"=cmax),trace=0)
hybridoptimalBIC2 = step(c0,direction="both",
scope=list("lower"=c0,"upper"=cmax),trace=0,k=log(3044))
```

We see from the summary output that the majority of the coefficients are significant with the departures from this coming from the State factor variable for certain states. In the AIC model we see that Poverty Percent is not significant with a p value of 0.0628, however we see in the BIC model this variable is dropped and as Poverty Percent will have a high correlation with Percent Unemployed 16 and Over we adopt the BIC model.

```{r statestepwise 2, include=FALSE}
# Produce summary and coefficients of state stepwise regression models
coef(hybridoptimalAIC2)
coef(hybridoptimalBIC2)
summary(hybridoptimalAIC2)
summary(hybridoptimalBIC2)
```

Below we can see the residual and QQplot for the BIC model. In the residual plot on the left hand side we see constant variance and a mean of zero with the majority of points near the horizontal red line. This is similar to what we saw when we didn't include the states in the selection process. Similarly, we see in the QQplot the majority of points on or near the reference line suggesting that the errors are indeed normally distributed, again as we see in the above case of not including states in the selection process.

```{r statestepwise 3, echo=FALSE, fig.asp=0.5, fig.cap="Residuals plot and QQ plot of hybrid stepwise model with states using BIC"}
# Residual plots of state stepwise regression
par(mfrow=c(1,2))
plot(hybridoptimalBIC2, which=c(1,2))
```

In order to test whether the addition of states in the model makes a significant difference on the model in the stepwise selection process we perform an F-Test using the anova function. Computing this we get a p-value of less than $2.2x10^-16$, thus giving us strong evidence that we should reject the null hypothesis that the model not including states is better. Hence, the BIC model including the states as a factor variable is a better fit of the data according to the anova function.

```{r statestepwise 4, include=FALSE}
# Perform F-Test on the BIC models above
anova(hybridoptimalBIC,hybridoptimalBIC2)
```

To complete the section on stepwise regression we perform leave one out cross validation in order to compute the $R^2$ of the stepwise models and also the Root Mean Squared Error. This allows us to compare these models with the Ridge and Lasso models we generate in the enxt sections. Performing this we see the AIC has an $R^2$ value of 0.4629, whereas in the BIC model we have an $R^2$ value of 0.4621.

```{r LOOCVforStepwise, cache=TRUE, include=FALSE}
library(caret)
#specify the cross-validation method
#fit a regression model and use LOOCV to evaluate performance
loocv <- function(lm1, data=cancer) {
  ctrl <- trainControl(method = "LOOCV")
  xnam <- names(lm1$coefficients)[-1]
  fmla <- as.formula(paste("deathRate ~ ", paste(xnam, collapse= "+")))
  model <- train(fmla, data = data, method = "lm", trControl = ctrl)
  return(model)
}
hybridoptimalAIC.loocv <- loocv(hybridoptimalAIC, data=cancermodel)
hybridoptimalBIC.loocv <- loocv(hybridoptimalBIC,data=cancermodel)
```
### RIDGE Regression

To address the multi-collinearity present in the data we assess the viability of a RIDGE Regression Model. For this we fit all continuous variables as predictors using the `glmnet` library. 

The trace shows that the ridge penalisation method reduces many variables close to 0 but does not remove any variables from the model itself.

```{r, include=FALSE}
library(glmnet)

# Create Data Matrix
cancer.dm <- cancer %>% 
  select(!c("Geography", "medIncome", "binnedInc", "PctBlack", "deathRate")) %>%
  data.matrix()

lm.ridge <- glmnet(cancer.dm, cancer$deathRate, alpha=0)
lm.ridge.cv <- cv.glmnet(cancer.dm, cancer$deathRate, alpha=0)

# Create Model Plot Func
traceLogLambda <- function(lm1, lm1.cv, ylim=NULL, sub=NULL) {
  plot(lm1,"lambda",label = T, ylim=ylim)
  abline(v=log(lm1.cv$lambda.1se),col="red")
  abline(v=log(lm1.cv$lambda.min),col="blue")
  legend("bottomright",legend=c("Minimum lambda", "1 standard error larger lambda"),lty=c(1,1),col=c("blue","red"), ins=0.05)
  title(sub=sub)
  
}

traceLogLambda(lm.ridge, lm.ridge.cv, ylim=c(-20, 1), sub="RIDGE trace against log(lambda)")

lm.ridge.1se <- glmnet(cancer.dm, cancer$deathRate, alpha = 0, 
               lambda = lm.ridge.cv$lambda.1se)


lm.ridgefitted <- predict(lm.ridge.1se, newx=cancer.dm)
scatter.smooth(cancer$deathRate - lm.ridgefitted, x=lm.ridgefitted,
               xlab="Fitted", ylab="Residuals", 
               sub="Residuals vs Fitted for RIDGE Model")

```

Fitting a ridge model and using the value of $\lambda$ one standard error further away from the minimum does not remove any of the terms from the model and gives the following coefficients. We also include the parameter estimates when the parameters are scaled to highlight how significant the parameter is.


```{r, include=FALSE, cache=TRUE}
library(parallel)
library(foreach)
library(doParallel)
numCores <- detectCores()
registerDoParallel(numCores) 

n <- dim(cancer)[1]
dev.ratios <- rep(NA, n)
dev.ratios1 <- rep(NA, n)
lm.ridge.cv <- cv.glmnet(cancer.dm, cancer$deathRate, alpha=0, nfolds=n, parallel=TRUE)

lm.ridge.cv$lambda.1se

for (i in 1:n) {
  lm.ridge.1se <- glmnet(cancer.dm[-i,], cancer$deathRate[-i], alpha = 0, 
               lambda = lm.ridge.cv$lambda.1se)
  
  
  dev.ratios1[i] <- lm.ridge.1se$dev.ratio
  # lm.residuals <- predict(lm.ridge.1se, newx=cancer.dm[-i, ]) - cancer$deathRate[-i]
  # dev.ratios[i] <- 1 - sum(lm.residuals^2)/sum((cancer$deathRate[-i] -mean(cancer$deathRate[-i]))^2) 
  
}
mean(dev.ratios1)

```

\scriptsize

```{r, echo=FALSE}
library(knitr)


a <- data.frame(sapply(round(coef(lm.ridge.1se), 3), FUN=identity))
colnames(a) <- "Parameter Estimates"

# cancer.dm.scale <- cancer %>%
#   select(!c("Geography", "medIncome", "binnedInc", "PctBlack", "deathRate")) %>%
#   scale %>%
#   data.matrix()
# lm.ridge.scale.cv <- cv.glmnet(cancer.dm.scale, scale(cancer$deathRate), alpha = 0)
# lm.ridge.1se.scale <- glmnet(cancer.dm.scale, scale(cancer$deathRate), alpha = 0, lambda = lm.ridge.scale.cv$lambda.1se)
# a$"Scaled Parameter Estimates" <-sapply(round(coef(lm.ridge.1se.scale), 3), FUN=identity)

rownames(a) <- rownames(round(coef(lm.ridge.1se), 3))
shortNames <- c("Intercept", "IncidRT", "PctPov", "MedMale", "MedFemale", 
                "AvgHH", "PctMarr", "PctEmp", "PctUnemp", "PrivCov", "PubCov", 
                "MarrHH", "lPctBlack", "lMedInc")
colnames(a) <- colnames(shortNames)
a
kable(t(a))
```
\normalsize

Performing a leave one out cross validation gives an $R^2$ statistic of $0.4438$. Which is similar to the previous step wise regression method and is harder to diagnose therefore we conclude that it it is not a suitable model.

## Lasso Regression 

First we removed Geography because it is an id variable. We also removed binnedInc because it is measuring the same thing as medIncome.
We used the log-transformed PctBlack and medIncome to improve homoscedasticity and normality.

```{r Lasso data cleaning,echo=FALSE}
# cancer.lasso.dm <- cancer %>%
#   select(!c("Geography", "medIncome", "binnedInc", "PctBlack", "deathRate","MedianAgeFemale","PctMarriedHouseholds")) %>%
#   data.matrix()
cancer.lasso.dm <- cancer %>%
  select(!c("Geography", "medIncome", "binnedInc", "PctBlack", "deathRate")) %>%
  data.matrix()
```

We used cross-validation from the glmnet library which leaves out a 10th of the data every time.
We produced the following plot of mean-squared error against log(lambda). We decided to use the 1-standard-error-lambda because this likely to shrink some predictor variables to zero, performing variable selection. We prefer a simpler model.

```{r Lasso cv, echo=FALSE}
set.seed((934))
lm.lasso <-glmnet(cancer.lasso.dm,cancer$deathRate,alpha=1)
lm.lasso.cv <- cv.glmnet(cancer.lasso.dm,cancer$deathRate,alpha=1)
plot(lm.lasso.cv)
abline(v=log(lm.lasso.cv$lambda.1se),col="red")
abline(v=log(lm.lasso.cv$lambda.min),col="blue")
legend("topright",legend=c("Minimum lambda", "1 standard error larger lambda"),lty=c(1,1),col=c("blue","red"), ins=0.1)
lm.lasso.cv$lambda.1se
lm.lasso.1se <- glmnet(cancer.lasso.dm,cancer$deathRate,lambda = lm.lasso.cv$lambda.1se,alpha=1)

lm.lasso.1se.fitted <- predict(lm.lasso.1se, newx=cancer.lasso.dm)
lm.lasso.1se.residuals <- cancer$deathRate - lm.lasso.1se.fitted

```
We produced a plot that shows a trace of each parameter estimate for different values of log(lambda).
We scaled the model for a better visual interpretation.

```{r Lasso trace, echo=FALSE}
lm.lasso.scaled<-glmnet(scale(cancer.lasso.dm),cancer$deathRate,alpha=1)
traceLogLambda(lm.lasso.scaled,lm.lasso.cv,ylim=c(-10,10), sub="LASSO trace against log(lambda)")
```
The Lasso regression using 1-standard-error-lambda produces the following parameter estimates.

\scriptsize
```{r Lasso estimates,echo=FALSE}
b <- data.frame(sapply(round(coef(lm.lasso.1se), 3), FUN=identity))
colnames(b) <- "Parameter Estimates"
rownames(b) <- rownames(round(coef(lm.lasso.1se), 3))
kable(b)
```
\normalsize

### Interpretation

From the estimates output above, the Lasso regression eliminates a considerable number of predictor variables. The Lasso regression model has the following non-zero predictor estimates: __Incidence Rate__, __Percent Employed 16 and Over__, __Percent Private Coverage__,  __Percent Employer Provided Private Coverage__, __Education Levels__, __Log of Percent Black__,
__Log of Median Income__.

$$
Death Rate = 470.47 + 0.23\ IncidenceRate -0.25\ Pecent Employed\_Over16 + 0.29\ PercentUnemployed\_Over16 \\- 0.32\  PercentPrivateCoverage  +0.21\ Percent Employer Provided Private Coverage \\-10.05\ Education Levels +0.90\ Log PercentBlack
-32.39\ Log Median Income
$$
The two age variables and Average Household Size are removed from the model. This agrees with our EDA which showed that they have very close to zero correlation coefficient with Death Rate.
Both employment variables and two out of three healthcare coverage variables are included in this model. We believe this is reasonable because although they showed evidence of collinearity/multicollinearity, we did not have strong arguments to remove any of them. Therefore, we agreed with the variable selection suggested by the Lasso Regression.

Most predictor estimates are smaller than one. However the estimates for Education Levels and Log Median Income are -10.05 and -32.39 respectively which are exceptionally large and negative. This means that they have much larger impact on the predicted death rates than other predictor variables. It makes sense that Median Income has the largest impact because patients in counties with higher median income tend to be able to afford better treatment which reduce mortality rates.

We calculated the R-squared using Leave-One-Out Cross Validation. The R-squared for this Lasso Regression model is 0.4441.
 
```{r LASSO_RMSE, include=FALSE, cache=TRUE}
numCores <- detectCores()
registerDoParallel(numCores) 

n <- dim(cancer)[1]
lasso.dev.ratios <- rep(NA, n)
lasso.dev.ratios1 <- rep(NA, n)
lm.lasso.cv <- cv.glmnet(cancer.lasso.dm, cancer$deathRate, alpha=0, nfolds=n, parallel=TRUE)

lm.lasso.cv$lambda.1se

for (i in 1:n) {
  lm.lasso.1se <- glmnet(cancer.dm[-i,], cancer$deathRate[-i], alpha = 0, 
               lambda = lm.lasso.cv$lambda.1se)
  
  
  lasso.dev.ratios1[i] <- lm.ridge.1se$dev.ratio
  # lm.residuals <- predict(lm.ridge.1se, newx=cancer.dm[-i, ]) - cancer$deathRate[-i]
  # dev.ratios[i] <- 1 - sum(lm.residuals^2)/sum((cancer$deathRate[-i] -mean(cancer$deathRate[-i]))^2) 
  
}
mean(lasso.dev.ratios1)

for (i in 1:n) {
  lm.lasso.min <- glmnet(cancer.dm[-i,], cancer$deathRate[-i], alpha = 0, 
               lambda = lm.lasso.cv$lambda.min)
  
  
  lasso.dev.ratios1[i] <- lm.ridge.1se$dev.ratio
  # lm.residuals <- predict(lm.ridge.1se, newx=cancer.dm[-i, ]) - cancer$deathRate[-i]
  # dev.ratios[i] <- 1 - sum(lm.residuals^2)/sum((cancer$deathRate[-i] -mean(cancer$deathRate[-i]))^2) 
  
}
mean(lasso.dev.ratios1)



```


## Statistical Interpretation and Validation

```{r ResidualPlot}

# So wanna make a plot of residuals vs fitteds
library(ggplot2)

multResidualPlot <- function(residual.list, fitted.list, models) {
  
  
  # a <- data.frame(x=fitted.list[[1]], y=residual.list[[1]], col=rep('blue', times=length(fitted.list[[1]])))
  a <- data.frame()
  for (i in 1:length(models)) {
    x <- fitted.list[[i]] 
    y <- residual.list[[i]]
    model <- rep(models[i], times=length(x))
    
    df.temp <- data.frame(x, y, model)
    colnames(df.temp) <- c("x", "y", "model")
    a <- rbind(a, df.temp)
  }
  mrp <- ggplot(a, aes(x=x, y=y, colour=model)) + 
    geom_point(alpha=0.3, size=0.75) + geom_smooth() +
    labs(x="Fitted Values", y="Residuals", title="Residuals vs Fitted Values", model="Models")
  return(mrp)
}

lm.ridge.residuals <- cancer$deathRate - lm.ridgefitted
residual.list <- list(
  hybridoptimalAIC$residuals,
  hybridoptimalBIC$residuals,
  lm.ridge.residuals,
  lm.lasso.1se.residuals
)

fitted.list <- list(
  hybridoptimalAIC$fitted.values,
  hybridoptimalBIC$fitted.values,
  lm.ridgefitted,
  lm.lasso.1se.fitted
)
mrp <- multResidualPlot(residual.list, fitted.list,
                 c("Hybrid Optimal AIC", "Hybrid Optimal BIC", "RIDGE", "LASSO"))
mrp
```
```{r, eval=FALSE, include=FALSE}
library(plotly)
### INTERACTIVE RESIDUAL
ggplotly(mrp)
```

Produce map of each states average residual value for the LASSO model.

```{r, include=FALSE}
#average for each state
averageresidual=round(tapply(lm.lasso.1se.residuals,cancermodel2$State,mean)[-c(2,12)],0)
states <- map_data("state")
averagedf <- data.frame(region=unique(states$region), averageresidual)
mergedf <- merge(states, averagedf, by="region")
statenames <- data.frame(region=tolower(state.name), clong=state.center$x, clat=state.center$y)
statenames <- merge(statenames, averagedf, by="region")
statenames$lab <- paste(statenames$region, '\n', statenames$averageresidual, sep="")
```



```{R USMAP, echo=FALSE, cache=TRUE}
qplot(long, lat, data=mergedf, geom="polygon", fill=averageresidual, group=region) + 
  scale_fill_gradient(averageresidual,low="yellow",high="red") +
  geom_text(data=statenames,aes(clong,clat,label=averageresidual,inherit.aes = FALSE,label.size=0.001)) +
  theme(axis.line=element_blank(),axis.text.x=element_blank(),
        axis.text.y=element_blank(),axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(), legend.title = element_blank(),
        panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),plot.background=element_blank(),plot.title =element_text(hjust=0.5) ) +
    ggtitle("Heatmap of Average Residuals of US States") 
```

# References 

<div id="refs"></div>

# Appendix

```{r get-labels, echo = FALSE}
labs = all_labels()
labs = setdiff(labs, c("setup", "get-labels"))
```

```{r all-code, ref.label=labs, eval=FALSE}
```
  
