---
title: "ST404 Assignment 2"
author: "Frank Or, Remos Gong, Sam Glanfield, Thomas Broadbent"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2: default
  pdf_document: 
    number_sections: true
fontsize: 11pt
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bookdown)
library(knitr)
```

```{r libraries, include=FALSE}
library(dplyr)
library(car)
```

\pagebreak 

# Findings

# Statistical Methodology

```{r loadData, include=FALSE}
## Included Libraries
# For pipe operator and general mutation
load('cancer.rdata')

```

We first combined our results and findings from our preliminary EDA which is mainly discussed in [Outliers](#outliers) and [Transformations](#transformations).

## Outliers {#outliers}

Counties with high Incidence Rates, namely 'Union County, Florida' and 'Williamsburg City, Virginia'. We looked into the cook's distance plots and noticed only 'Williamsburg City, Virginia' has large cook's distance and hence influential. The first cook's distance plot used a linear model with only incidenceRate as the predictor variable. The second used all the numerical variables. We concluded although 'Union County, Florida' has high leverage, it is not influential and hence should be kept in our data set.

```{r incidenceRateRemoval, echo=FALSE}
# Cook's distance Plot
par(mfrow=c(1,2))
plot(lm(deathRate ~ incidenceRate,data=cancer),4)
plot(lm(deathRate ~ .,data=cancer[-c(1,4)]),4)
# Removing outlier incidence rates 'Williamsburg City, Virginia'
cancer <- filter(cancer, incidenceRate <= 1100)
```

For counties with Average Household Size less than one we took the decision to scale the transformations by 100 and keep them in the dataset. This fixed the normality of AvgHouseholdSize as shown in the histogram.

```{r averageHouseholdSize, include=FALSE}
# Scale average household sizes that are less than 1 by 100
cancer[which(cancer$AvgHouseholdSize < 1), ]$AvgHouseholdSize <- 
  100*cancer[which(cancer$AvgHouseholdSize < 1), ]$AvgHouseholdSize
hist(cancer$AvgHouseholdSize, breaks=30, xlab="AvgHouseholdSize", main="Histogram of AvgHouseholdSize")
```

We also see counties with missing values in Percentage Employed 16 and Over and we conclude that the data is Missing Completely at Random. In order to rectify this we impute this data by fitting a linear regression model of Percentage Employed 16 and Over on the remaining variables to estimate what these values would be.

```{r impute, include=FALSE}
# Impute the missing data seen in the dataset
mod1=lm(PctEmployed16_Over~+deathRate+incidenceRate+medIncome+binnedInc+povertyPercent+MedianAgeMale+MedianAgeFemale+AvgHouseholdSize+PercentMarried+PctUnemployed16_Over+PctPrivateCoverage+PctEmpPrivCoverage+PctPublicCoverage+PctBlack+PctMarriedHouseholds+Edu18_24,cancer)
missdf = cancer[which(is.na(cancer$PctEmployed16_Over)==TRUE),]
imputed = predict(mod1,missdf)
cancer[which(is.na(cancer$PctEmployed16_Over)==TRUE),"PctEmployed16_Over"] = imputed

```


## Transformations {#transformations}

We transform Percent Black by first shifting the values upwards by 0.05, to ensure we have no zero values, then take a log transform. We also transform the Median Income by again taking a log trasformation. We do these transformations to ensure the data is not heavily skewed and allow for a more accurate model.

```{r transformations, include=FALSE}
# Log transforming the heavily skewed distributions of PctBlack and medIncome
cancer$logpctblack = log(cancer$PctBlack+0.05)
cancer$logmedincome = log(cancer$medIncome)
```

The following residual plots show the improvements in homoscedasticity in PctBlack and medIncome after log-transform respectively.

```{r heteroscedasticity1,echo=FALSE}
# Showing improvements in homoscedasticity in PctBlack and medIncome
par(mfrow=c(1,2))
plot(lm(deathRate~PctBlack,data=cancer),1)
plot(lm(deathRate~logpctblack,data=cancer),1)
plot(lm(deathRate~medIncome,data=cancer),1)
plot(lm(deathRate~logmedincome,data=cancer),1)
```

We also notice slight evidence of heteroscedasticity in PctEmpPrivCoverage as shown in the residual plot below.
However, since it is not severe and the suggested power transformation using spread level plot makes it more heteroscedastic, we decided not to perform transformation on PctEmpPrivCoverage.

```{r Spreadlevelplot, fig.show='hide'}
# Finding power transformation for PctEmpPrivCoverage
spreadLevelPlot(lm(deathRate~PctEmpPrivCoverage, data=cancer))
```


```{r heteroscedasticity2,echo=FALSE}
plot(lm(deathRate~PctEmpPrivCoverage,data=cancer),1)
```



## Modelling Approach and Variable Selection

### AIC Forward and Backward Variable Selection

### RIDGE Regression

## Statistical Interpretation and Validation

# References 

<div id="refs"></div>

# Appendix

```{r get-labels, echo = FALSE}
labs = all_labels()
labs = setdiff(labs, c("setup", "get-labels"))
```

```{r all-code, ref.label=labs, eval=FALSE}
```
  
